Timeliness of security updates for Android

### 2014-10-22 securityupdates.tex
Add more data on analysis of external projects

This data shows a large gap in the middle where we do not know of any discoveries of root equivalent vulnerabilities affecting all of Android.
The cause of this quiet period is unclear, perhaps from Figure~\ref{fig:proportioninsecure} the reason was that since most devices were exposed to known vulnerabilities there was no point in looking for new ones?
Or that manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices?
Or that manufacturer specific vulnerabilities proved to be easier to find and so people looking for vulnerabilities adjusted their focus?

\section{Android Ecosystem}
\subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
While we have not tried to exploit these vulnerabilities on the devices in order to test whether the devices are actually vulnerable, we are confident that most devices will be if they are running a vesion known to be vulnerable. \todo{WHY?}

\subsection{Android update process}
\label{sec:android_update_process}
\begin{figure}
 \centering
 \def\svgwidth{\columnwidth}
 \import{figures/}{update_ecosystem.pdf_tex}
 \caption{Flow of updates between participants in the Android ecosystem.
 Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.
 Dotted arrows indicate flows where we do not know how many updates are being produced as we can't measure those flows directly as they are not public.\todo{Update this diagram}}
 \label{fig:update_ecosystem}
\end{figure}

A model of the Android update process is shown in Figure~\ref{fig:update_ecosystem}.
There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and various Apache Java libraries and so can include any of the versions of those projects, including those which fix security vulnerabilities in them.
It also incorporates various drivers for different bits of hardware.
The Android platform is then built on top of those with kernel and userspace components by Google. \todo{ work out the exact arrangements.}
Usually the code for each update is kept secret\todo{Can we quantify this keeping the code secret? Is it worth it?} until after the update has been published though through the Open Handset Alliance.
Manufacturers in the  alliance receive advanced copies in order to prepare handsets.
Then the Manufacturers take the update and customise it before passing it on to the operator.
The Operator may then make further customisations and do further testing before shipping the update to the Device.
Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google are the same entity (at least for the purpose of a particular phone, such as with Nexus phones).
Sometimes Manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqlite in their release of Android
%An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
\todo{talk about update latency, quantify this in the text as per the figure, giving more explanation, use statistics from samsung-updates.com -> how many binaries are there per device? Do we know the variation between network operators?}

@@

\subsection{Monetisation}
\todo{redo this section in light of related work}

>>>>

This data shows a large gap in the middle where we do not know of any discoveries of root equivalent vulnerabilities affecting all of Android.
The cause of this quiet period is unclear, perhaps from Figure~\ref{fig:proportioninsecure} the reason was that since most devices were exposed to known vulnerabilities there was no point in looking for new ones?
Or that manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices?
Or that manufacturer specific vulnerabilities (which we are ignoring) proved to be easier to find and so people looking for vulnerabilities adjusted their focus?

\section{Android Ecosystem}
\subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
While we have not tried to exploit these vulnerabilities\todo{ethics} on the devices in order to test whether the devices are actually vulnerable, we are confident that most devices will be if they are running a vesion known to be vulnerable. \todo{WHY?}

\subsection{Android update process}
\todo{rewrite this}
\label{sec:android_update_process}
\begin{figure}
 \centering
 \def\svgwidth{\columnwidth}
 \import{figures/}{update_ecosystem.pdf_tex}
 \caption{Flow of updates between participants in the Android ecosystem.
 Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.
 Dotted arrows indicate flows where we do not know how many updates are being produced as we can't measure those flows directly as they are not public.\todo{Update this diagram}}
 \label{fig:update_ecosystem}
\end{figure}

A model of the Android update process is shown in Figure~\ref{fig:update_ecosystem}.
There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and so can include any of the versions of those projects, including those which fix security vulnerabilities in them.
There are \avoNumExternalProjects\ external projects in Android, contributing \avoTotalExternalLines\ lines of code.
We analysed the top \avoNumBigExternalProjects\ by lines of code (contributing \avoNumBigExternalLinesOfCode, \avoBigExternalLinesOfCodePerc\ of the total) and were able to automatically extract the versions of those projects included in different versions of Android for \avoNumBigExternalProjectsAnalysed.
We found \avoBigExternalTotalVersions\ distinct versions, a median of \avoBigExternalMedianVersions\ and mean of \avoBigExternalMeanVersions\ versions per project.
Android also incorporates various drivers for different bits of hardware.
The Android platform is then built on top of those with kernel and userspace components by Google. \todo{ work out the exact arrangements.}
Usually the code for each update is kept secret\todo{Can we quantify this keeping the code secret? Is it worth it?} until after the update has been published though through the Open Handset Alliance.
Manufacturers in the  alliance receive advanced copies in order to prepare handsets.
Then the Manufacturers take the update and customise it before passing it on to the operator.
The Operator may then make further customisations and do further testing before shipping the update to the Device.
Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google are the same entity (at least for the purpose of a particular phone, such as with Nexus phones).
Sometimes Manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example including a broken daily build of sqlite in their release of Android
%An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
\todo{talk about update latency, quantify this in the text as per the figure, giving more explanation, use statistics from samsung-updates.com -> how many binaries are there per device? Do we know the variation between network operators?}

@@

\subsection{Monetisation}
\todo{redo this section in light of related work and to fit into ecosystem discussion}
